Cartogram Mapping and its Application to Cancer Data Visualisation

### 2019-07-22 review_article.Rmd
Reported measures

<!-- Measures of cancer -->

Almost 100 years of cancer mapping in the United States and the United Kingdom has increased the effectiveness in the presentation of unbiased rates. Mortality rates are now often presented as relative rates of risk across the population, and age adjusted to correct for the the higher prevalence of cancers in older populations. Howe [-@HEDP] describes Stocks development of the standardised mortality ratios through the 1930s.

@@

### Reported measures in cancer atlases

The list below summarises these measures, and provides a definition of each measure.

@@

<!--
@DMAHP cites Cruickshank's (1947) discussion of using visuals as 'formal statistical assessment of the spatial pattern' as a major advancement.
@MACM also recognises the intuition that can be derived from choropleths must be 'validated by rigorous statistical analyses.'

@CD considers cartographic data analysis in practice. Presenting a clear guide for preparing cartograms for use with both qualitative and quantitative information.
The authors recognise that the creation of cartograms was largely the work of professional cartographers until the innovation of geographical information systems that welcomed map users as map creators. These systems are utilised depending on ‘the effectiveness, efficiency, and satisfaction of the map products (Nielsen 1994)’. 
-->

@@

<!-- Going from choropleth to cartogram 
A cartogram is a map purposely distorted so that its spatial properties represent quantities
not directly associated with position on the globe. As thematic maps, cartograms emphasize
the distribution of a variable by changing the area (or lengths) of objects on the map.
Our approach maintains continuity, and preserves
many local features of cartographic lines that provide visual clues to the identity of the distorted
objects.
@ACCAC
-->

@@

With that in mind, the intended user and message to communicate should drive map selection.
first is the nature or inherent properties of the device itself
second is the ease and accuracy with which map users are able to extract information
from it

@SAMGIS suggests it is the "investigators' objectives" that drive the "representation of diseases on maps". 
-infectious diseases may require choropleth methods

@@

# Animation

@@

What software did we use, for e.g.

@@

<!--
# Software
Commonly available tools to implement.
-->

@@

Driven by progressions in computing, statistical methods and collection of data.

Models summarise relationships between counts of diseases and other factors.

Recognition of biases when utilising the geography to map people. Small areas contain more people, work towards representing regions of importance more clearly.
-->

>>>>

<!-- Measures reported in cancer atlases -->

Almost 100 years of cancer mapping in the United States and the United Kingdom has seen increased effectiveness in the presentation of unbiased rates. Mortality rates are now often presented as relative rates of risk across the population, and age adjusted to correct for the the higher prevalence of cancers in older populations. Howe [-@HEDP] describes Stocks development of the standardised mortality ratios through the 1930s.

@@

The list below summarises the measures presented in published cancer atlases, and provides a definition of each measure.

@@

@DMAHP cites Cruickshank's (1947) discussion of using visuals as 'formal statistical assessment of the spatial pattern' as a major advancement.

When presenting cancer atlases, @MACM believes the intuition that can be derived must be 'validated by rigorous statistical analyses.' Additional statistics are often

@@

<!-- Going from choropleth to cartogram

A cartogram is an alternative map visual. 
An intention distortion is applied to a geographic map base,

A cartogram is a map purposely distorted so that its spatial properties represent quantities
not directly associated with position on the globe. As thematic maps, cartograms emphasize
the distribution of a variable by changing the area (or lengths) of objects on the map.
Our approach maintains continuity, and preserves
many local features of cartographic lines that provide visual clues to the identity of the distorted
objects.
@ACCAC

@CD considers cartographic data analysis in practice. Presenting a clear guide for preparing cartograms for use with both qualitative and quantitative information.
The authors recognise that the creation of cartograms was largely the work of professional cartographers until the innovation of geographical information systems that welcomed map users as map creators. These systems are utilised depending on ‘the effectiveness, efficiency, and satisfaction of the map products (Nielsen 1994)’.

@@

To choose an approporiate map display, the map creator must consider the intended user, and message the map will communicate.
It is the objectives of the investigator that will drive the choice of representation [@CPISACA].

There are two keys to drive the choice of display: the properties of the visualisation, and the ease or accuracy of information extraction for map users @SAMGIS.

@@

# Animation and Interactivity

Recent developments of technology allowed interactive

@@

@@

<!--
# Software
Commonly available tools to implement.
-->
@tidyerse for data analysis, @spData and @sf to implement plotting of spatial data.
@sugarbag to create hexagon tesselation. @ggthemes to enhance plot displays.
@cartogram for contiguous and non-contiguous cartogram displays.

@@
